Code for Breast Cancer Pain Study

### 2020-07-28 manuscript_code_20200728.Rmd
Blight added

colnames(patients_file)[which(colnames(patients_file)=="POVERTY")] = "Poverty"
colnames(patients_file)[which(colnames(patients_file)=="VACANCY")] = "Vacancy"
colnames(patients_file)[which(colnames(patients_file)=="WALKABILITY")] = "Walkability"
colnames(patients_file)[which(colnames(patients_file)=="UNEMPLOYMENT")] = "Unemployment"
colnames(pcm_file)[which(colnames(pcm_file)=="a_CCMA06")] = "Pain"
```

@@

* "joined_zipcode" file merged to get data about poverty, vacancy, unemployment levels for more zip codes.   
* Poverty data (from "joined_zipcode")

@@

dmd.pvt = dmd.pvt[-which(dmd.pvt$Poverty>70),]
dmd.pvt = dmd.pvt[-which((dmd.pvt$Vacancy>40) & (dmd.pvt$mean.Pain<2)),]
cat(paste("Three outliers excluded"),"\n")

@@

```{r}
final.data <- 
  cdmd.pvt %>%
  select("Pain",
         "Poverty","Vacancy",
         "Occupation","Marital.status","Insurance.type",
         "Cancer.stage","Chemo","Surgery",
         "Race","Age",
         "Metastasis","Histology",
         "Primary.Site.1","Primary.Site.2",
         "HER2","ER","PR")
```

@@

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")

@@

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

**_Vacancy variable_**:

@@

hist(cdmd.pvt$Vacancy,main="Vacancy")
cat(paste("Number of zipcode =",length(unique(cdmd.pvt$zip))),"\n")
cat(paste("Median of Vacancy =",round(median(cdmd.pvt$Vacancy),digits=3)),"\n")

@@

### Vacancy

* Pain ~ Vacancy

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Vacancy, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

meandata = data.frame(Vacancy = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
                      Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
                      zip = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$zip)
meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
# hist(round(meandata$Pain))
# summary(glm(round(meandata$Pain) ~ meandata$Vacancy,family = "poisson"))
# summary(glm.nb(round(meandata$Pain) ~ meandata$Vacancy))

@@

pdf(6,6,file=paste("Pain_Vacancy_plot.pdf"))
par(mar=c(4,4,2,2))
plot(x=meandata$Vacancy,y=meandata$Pain,
     xlim=c(0,60),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
     axes=F,
     xlab=NA,ylab=NA,
     cex=cex,lwd=2.5)
box()
abline(0, 0, lty=1, col="grey")
abline(v=0, lty=1, col="grey")
# axis(side=1, tck=-.015,labels=NA)
# axis(side=1, lwd=0, line=-1, cex=0.2,cex.axis=0.9)
axis(side=1, tck=-0.015,lwd=0,line=-.5,cex.axis=1.5)
axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
mtext(side=1, "Vacancy", line=2.5, cex=1.5)
mtext(side=2, "Pain average", line=2, cex=1.5)

@@

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Vacancy, data = cdmd.pvt, family = "poisson")
# summary(p.glm)
# summary(zeroinfl(Pain ~ Poverty, data = cdmd.pvt, dist="poisson"))
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

meandata = data.frame(Vacancy = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
                      Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
                      zip = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$zip)
meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
# hist(round(meandata$Pain))
# summary(glm(round(meandata$Pain) ~ meandata$Vacancy,family = "poisson"))
# summary(glm.nb(round(meandata$Pain) ~ meandata$Vacancy))

par(mfrow=c(1,2))
cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))
# pdf(6,6,file=paste("Pain_Vacancy_plot.pdf"))
plot(x=meandata$Vacancy,y=meandata$Pain,cex=cex,xlab="Vacancy",ylab="Pain average")
# plot(x=aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
#      y=aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
#      xlab="Vacancy",ylab=paste(pain.type,"Pain mean within Zip"))
x=seq(0,80,by=1)
points(x=x,y=fhat(x),type="l",lwd=2)
# dev.off()
```

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age")
explvars.2 = c("Vacancy","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age")
explvars = c("Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")
full.formula.1 = as.formula(paste(respvar," ~ ",paste(explvars.1,collapse=" + ")))
full.formula.2 = as.formula(paste(respvar," ~ ",paste(explvars.2,collapse=" + ")))
full.formula = as.formula(paste(respvar," ~ ",paste(explvars,collapse=" + ")))
```

@@

## Poisson count model
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
tablevars = c("(Intercept)","Poverty","Vacancy",unlist(vars.list)) # all variables that will be printed in the table

@@

explvars.2 = c("Vacancy","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars = c("Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

## Poisson count model:
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
tablevars = c("(Intercept)","Poverty","Vacancy",unlist(vars.list)) # all variables that will be printed in the table

>>>>

colnames(patients_file)[which(colnames(patients_file)=="POVERTY")] = "Poverty"
colnames(patients_file)[which(colnames(patients_file)=="VACANCY")] = "Blight"
colnames(patients_file)[which(colnames(patients_file)=="WALKABILITY")] = "Walkability"
colnames(patients_file)[which(colnames(patients_file)=="UNEMPLOYMENT")] = "Unemployment"
colnames(pcm_file)[which(colnames(pcm_file)=="a_CCMA06")] = "Pain"
```

@@

* "joined_zipcode" file merged to get data about poverty, Blight, unemployment levels for more zip codes.   
* Poverty data (from "joined_zipcode")

@@

dmd.pvt = dmd.pvt[-which(dmd.pvt$Poverty>70),]
dmd.pvt = dmd.pvt[-which((dmd.pvt$Blight>40) & (dmd.pvt$mean.Pain<2)),]
cat(paste("Three outliers excluded"),"\n")

@@

```{r}
final.data <- 
  cdmd.pvt %>%
  select("Pain",
         "Poverty","Blight",
         "Occupation","Marital.status","Insurance.type",
         "Cancer.stage","Chemo","Surgery",
         "Race","Age",
         "Metastasis","Histology",
         "Primary.Site.1","Primary.Site.2",
         "HER2","ER","PR")
```

@@

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")

@@

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

**_Blight variable_**:

@@

hist(cdmd.pvt$Blight,main="Blight")
cat(paste("Number of zipcode =",length(unique(cdmd.pvt$zip))),"\n")
cat(paste("Median of Blight =",round(median(cdmd.pvt$Blight),digits=3)),"\n")

@@

### Blight

* Pain ~ Blight

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Blight, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

meandata = data.frame(Blight = aggregate(Blight ~ zip,cdmd.pvt,mean)$Blight,
                      Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
                      zip = aggregate(Blight ~ zip,cdmd.pvt,mean)$zip)
meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
# hist(round(meandata$Pain))
# summary(glm(round(meandata$Pain) ~ meandata$Blight,family = "poisson"))
# summary(glm.nb(round(meandata$Pain) ~ meandata$Blight))

@@

pdf(6,6,file=paste("Pain_Blight_plot.pdf"))
par(mar=c(4,4,2,2))
plot(x=meandata$Blight,y=meandata$Pain,
     xlim=c(0,60),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
     axes=F,
     xlab=NA,ylab=NA,
     cex=cex,lwd=2.5)
box()
abline(0, 0, lty=1, col="grey")
abline(v=0, lty=1, col="grey")
# axis(side=1, tck=-.015,labels=NA)
# axis(side=1, lwd=0, line=-1, cex=0.2,cex.axis=0.9)
axis(side=1, tck=-0.015,lwd=0,line=-.5,cex.axis=1.5)
axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
mtext(side=1, "Blight", line=2.5, cex=1.5)
mtext(side=2, "Pain average", line=2, cex=1.5)

@@

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Blight, data = cdmd.pvt, family = "poisson")
# summary(p.glm)
# summary(zeroinfl(Pain ~ Poverty, data = cdmd.pvt, dist="poisson"))
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

meandata = data.frame(Blight = aggregate(Blight ~ zip,cdmd.pvt,mean)$Blight,
                      Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
                      zip = aggregate(Blight ~ zip,cdmd.pvt,mean)$zip)
meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
# hist(round(meandata$Pain))
# summary(glm(round(meandata$Pain) ~ meandata$Blight,family = "poisson"))
# summary(glm.nb(round(meandata$Pain) ~ meandata$Blight))

par(mfrow=c(1,2))
cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))
# pdf(6,6,file=paste("Pain_Blight_plot.pdf"))
plot(x=meandata$Blight,y=meandata$Pain,cex=cex,xlab="Blight",ylab="Pain average")
# plot(x=aggregate(Blight ~ zip,cdmd.pvt,mean)$Blight,
#      y=aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
#      xlab="Blight",ylab=paste(pain.type,"Pain mean within Zip"))
x=seq(0,80,by=1)
points(x=x,y=fhat(x),type="l",lwd=2)
# dev.off()
```

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age")
explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age")
explvars = c("Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")
full.formula.1 = as.formula(paste(respvar," ~ ",paste(explvars.1,collapse=" + ")))
full.formula.2 = as.formula(paste(respvar," ~ ",paste(explvars.2,collapse=" + ")))
full.formula = as.formula(paste(respvar," ~ ",paste(explvars,collapse=" + ")))
```

@@

## Poisson count model
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
tablevars = c("(Intercept)","Poverty","Blight",unlist(vars.list)) # all variables that will be printed in the table

@@

explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars = c("Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

## Poisson count model:
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
tablevars = c("(Intercept)","Poverty","Blight",unlist(vars.list)) # all variables that will be printed in the table
